From physical principles to ???

### 2015-12-14 2015CompleteDerivation/completeDerivation.tex
Using mathcal for state spaces

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}

@@

\begin{defn}\label{reducible_state_space}
The state space $\mathbbm{C}$ for a homogeneous decomposable system is an additive abelian (i.e. commutative) group.
\end{defn}

\begin{justification}
We claim $\mathbbm{C}$ is an additive monoid. There exist a law of composition $+ : \mathbbm{C} \times \mathbbm{C} \rightarrow \mathbbm{C}$ that takes two states and returns one that is the physical composition of the two. The domain and codomain match because the system is homogeneous. The law is commutative $\mathbbm{c}_1 +\mathbbm{c}_2 = \mathbbm{c}_2+\mathbbm{c}_1$ and associative $(\mathbbm{c}_1 + \mathbbm{c}_2) + \mathbbm{c}_3 = \mathbbm{c}_1 + (\mathbbm{c}_2 + \mathbbm{c}_3)$, as it does not matter in what order we physically compose the parts. There exist a unique zero element $\mathbbm{c} + 0 = \mathbbm{c}$ and it represent the physically empty state (i.e. no system).

We claim $\mathbbm{C}$ is an additive group. As we want to describe changes during the evolution (i.e. $\hat{\mathbbm{c}} = \mathbbm{c} +\Delta \mathbbm{c}$) we introduce an inverse $\mathbbm{c}_2$ for each element $\mathbbm{c}_1$ such that $\mathbbm{c}_1 + \mathbbm{c}_2 = 0$. The inverse may not represent a proper physical state, yet we'll still call $\mathbbm{C}$ the state space committing an abuse of terminology. The inverse $- : \mathbbm{C} \rightarrow \mathbbm{C}$ is physically unique and well defined.
\end{justification}

@@

We will call a classical system one that is infinitesimally reducible. That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system is equivalent to the sum of the states of the parts. Given the state space $\mathbbm{S}$ for the parts, giving a composite state $\mathbbm{c}$ is equivalent to give the amount of particles that are present in each state. Since the system is infinitesimally reducible, such amount can be increased or decreased infinitesimally and will be a real number. In other words: the classical state is a real valued distribution over $\mathbbm{S}$. This intuitive picture is unfortunately not suitable to be generalized to different contexts, so we develop another.

Consider the following expression:
\begin{align*}
\mathbbm{c} = \sum\limits_{i \in \mathbbm{I}} a_i \mathbbm{e}(i)
\end{align*}
$\mathbbm{I}$ is best description of an infinitesimal part that a composite system can give. $\mathbbm{e}(i)$ represents a composite system made of a unit amount of particles, all prepared in the state determined by $i$. $a_i$ represents a transformation that changes the whole state, without affecting the value $i$ for any infinitesimal part. For a classical system, $i$ represents the full information about the infinitesimal part, and therefore includes the value of all state variables; the only transformation left is increasing/decreasing the amount of particles by a scalar multiple. For a different system, where $i$ does not represent the full state of the parts, $\mathbbm{e}(i)$ will be an ensemble and $a_i$ will include internal transformations. This is intuitively how we decouple the composite state into the part $i$ that describes the infinitesimal parts, and the part $a_i$ that describes the rest. Giving a full description of the state space of a decomposable system means fully characterizing these components. And we can see how this fits for the classical case as well: the state variables gives us the description for an infinitesimal part, but the amount of particles for each state is a description of the composite state.

\begin{defn}\label{classical_vector space}
The state space $\mathbbm{C}$ for a homogeneous classical (i.e. infinitesimally reducible) system is a vector space over $\mathbbm{R}$. It admits a basis isomorphic to the state space $\mathbbm{S}$ of its particles (i.e. infinitesimal parts).
\end{defn}

@@

We now claim that the state space $\mathbbm{C}$ is a vector space over $\mathbbm{R}$. The abelian group $\mathbbm{C}$ can be extended with the operations defined by $A$, as each element $a \in A$ is a map $a : \mathbbm{C} \rightarrow \mathbbm{C}$. The map has the following properties: $(a_1 + a_2) \mathbbm{c} = a_1 \mathbbm{c} + a_2 \mathbbm{c}$, increasing the size of the system by the sum of two constant is the same as combining the separate increases, and $a (\mathbbm{c}_1 + \mathbbm{c}_2) = a \mathbbm{c}_1 + a \mathbbm{c}_2$, increasing the size of the total system is the same as the combination of the increased parts. $\mathbbm{C}$ is a module over $A$, which is a field and isomorphic to $\mathbbm{R}$.

Consider now the injection map $\mathbbm{e}: \mathbbm{S} \hookrightarrow \mathbbm{C}$ that for each particle state returns a composite state constituted only by a predetermined amount of particles in such state. The image of such injection is a basis for $\mathbbm{C}$. We first claim that all elements are linearly independent. Consider $\sum\limits_{\mathbbm{s}_i \in \mathbbm{S}} a_i \mathbbm{e}(\mathbbm{s}_i)$, with $a_i \in \mathbbm{R}$ and . This corresponds to a physical state with $a_i$ particles in each state $\mathbbm{s}_i$. This is equal to the empty state if and only if $a_i=0 \forall i$. This holds even if $\mathbbm{S}$ is not finite. We now claim they span. Consider $\mathbbm{c} + \sum\limits_{\mathbbm{s}_i \in \mathbbm{S}} a_i \mathbbm{e}(\mathbbm{s}_i)$, $a_i \in \mathbbm{R}$. This can always be made equal to the empty state by setting $a_i$ to be the opposite of the amount of particles of $\mathbbm{c}$ that are in state $\mathbbm{s}_i$.
\end{justification}

>>>>

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{dutchcal}

@@

\begin{defn}\label{reducible_state_space}
The state space $\mathcal{C}$ for a homogeneous decomposable system is an additive abelian (i.e. commutative) group.
\end{defn}

\begin{justification}
We claim $\mathcal{C}$ is an additive monoid. There exist a law of composition $+ : \mathcal{C} \times \mathcal{C} \rightarrow \mathcal{C}$ that takes two states and returns one that is the physical composition of the two. The domain and codomain match because the system is homogeneous. The law is commutative $\mathcal{c}_1 +\mathcal{c}_2 = \mathcal{c}_2+\mathcal{c}_1$ and associative $(\mathcal{c}_1 + \mathcal{c}_2) + \mathcal{c}_3 = \mathcal{c}_1 + (\mathcal{c}_2 + \mathcal{c}_3)$, as it does not matter in what order we physically compose the parts. There exist a unique zero element $\mathcal{c} + 0 = \mathcal{c}$ and it represent the physically empty state (i.e. no system).

We claim $\mathcal{C}$ is an additive group. As we want to describe changes during the evolution (i.e. $\hat{\mathcal{c}} = \mathcal{c} +\Delta \mathcal{c}$) we introduce an inverse $\mathcal{c}_2$ for each element $\mathcal{c}_1$ such that $\mathcal{c}_1 + \mathcal{c}_2 = 0$. The inverse may not represent a proper physical state, yet we'll still call $\mathcal{C}$ the state space committing an abuse of terminology. The inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ is physically unique and well defined.
\end{justification}

@@

We will call a classical system one that is infinitesimally reducible. That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system is equivalent to the sum of the states of the parts. Given the state space $\mathcal{S}$ for the parts, giving a composite state $\mathcal{c}$ is equivalent to give the amount of particles that are present in each state. Since the system is infinitesimally reducible, such amount can be increased or decreased infinitesimally and will be a real number. In other words: the classical state is a real valued distribution over $\mathcal{S}$. This intuitive picture is unfortunately not suitable to be generalized to different contexts, so we develop another.

Consider the following expression:
\begin{align*}
\mathcal{c} = \sum\limits_{i \in \mathbbm{I}} a_i \mathcal{e}(i)
\end{align*}
$\mathbbm{I}$ is best description of an infinitesimal part that a composite system can give. $\mathcal{e}(i)$ represents a composite system made of a unit amount of particles, all prepared in the state determined by $i$. $a_i$ represents a transformation that changes the whole state, without affecting the value $i$ for any infinitesimal part. For a classical system, $i$ represents the full information about the infinitesimal part, and therefore includes the value of all state variables; the only transformation left is increasing/decreasing the amount of particles by a scalar multiple. For a different system, where $i$ does not represent the full state of the parts, $\mathcal{e}(i)$ will be an ensemble and $a_i$ will include internal transformations. This is intuitively how we decouple the composite state into the part $i$ that describes the infinitesimal parts, and the part $a_i$ that describes the rest. Giving a full description of the state space of a decomposable system means fully characterizing these components. And we can see how this fits for the classical case as well: the state variables gives us the description for an infinitesimal part, but the amount of particles for each state is a description of the composite state.

\begin{defn}\label{classical_vector space}
The state space $\mathcal{C}$ for a homogeneous classical (i.e. infinitesimally reducible) system is a vector space over $\mathbbm{R}$. It admits a basis isomorphic to the state space $\mathcal{S}$ of its particles (i.e. infinitesimal parts).
\end{defn}

@@

We now claim that the state space $\mathcal{C}$ is a vector space over $\mathbbm{R}$. The abelian group $\mathcal{C}$ can be extended with the operations defined by $A$, as each element $a \in A$ is a map $a : \mathcal{C} \rightarrow \mathcal{C}$. The map has the following properties: $(a_1 + a_2) \mathcal{c} = a_1 \mathcal{c} + a_2 \mathcal{c}$, increasing the size of the system by the sum of two constant is the same as combining the separate increases, and $a (\mathcal{c}_1 + \mathcal{c}_2) = a \mathcal{c}_1 + a \mathcal{c}_2$, increasing the size of the total system is the same as the combination of the increased parts. $\mathcal{C}$ is a module over $A$, which is a field and isomorphic to $\mathbbm{R}$.

Consider now the injection map $\mathcal{e}: \mathcal{S} \hookrightarrow \mathcal{C}$ that for each particle state returns a composite state constituted only by a predetermined amount of particles in such state. The image of such injection is a basis for $\mathcal{C}$. We first claim that all elements are linearly independent. Consider $\sum\limits_{\mathcal{s}_i \in \mathcal{S}} a_i \mathcal{e}(\mathcal{s}_i)$, with $a_i \in \mathbbm{R}$ and . This corresponds to a physical state with $a_i$ particles in each state $\mathcal{s}_i$. This is equal to the empty state if and only if $a_i=0 \forall i$. This holds even if $\mathcal{S}$ is not finite. We now claim they span. Consider $\mathcal{c} + \sum\limits_{\mathcal{s}_i \in \mathcal{S}} a_i \mathcal{e}(\mathcal{s}_i)$, $a_i \in \mathbbm{R}$. This can always be made equal to the empty state by setting $a_i$ to be the opposite of the amount of particles of $\mathcal{c}$ that are in state $\mathcal{s}_i$.
\end{justification}
